% There is currently a problem with spacing somewhere so that Table of
% Contents, List of Tables, and List of Figures have the wrong amount
% of space.  Others are OK though...
\chapter*{Acknowledgements}
\addcontentsline{toc}{chapter}{Acknowledgements}


For the last eight years Trinity has been my academic home in the big smoke where I have been lucky enough to have met and received the help and kindness of so many people, without which I certainly would not have gotten to the brink of the big Dr. For even allowing me into Zoology I would like to thank both Prof. Trevor Hodkinson and Dr John Rochford who helped me transfer to Zoology in my first year (the second best decision of my life so far), saving me from the terrible clutches of Physics. Once safe within the interior of the Zoology Building I got to be part of many inspiring lectures including from Prof. Celia Holland who I thank in particular for inspiring me to chase my interests as an undergraduate and for guiding me so patiently during my final year thesis. A big thank you to my supervisor Dr Andrew Jackson who not only gave me the opportunity to continue to do science in Trinity but gave me the full license to do the science that put a smile on my face for the last four years and for providing so much support throughout. I have the Jackson lab to also thank for making my PhD so enjoyable, Luke McNally and Mafalda Viana helped immensely in guiding me through the art of science before back-up in the form of Adam Kane and his endless depravity for puns helped keep my heady morning cynicism at bay. The office would also certainly have been a lesser place without the Cooper due of Sive Finlay and Thomas Guillerme who along with Dr Natalie Cooper brought so much to the department both academically and personally. The door to our office was usually open and for good reason, the input, support and chin-wags from Deirdre McClean, Se{\'a}n Kelly and the whole cast of NERD club has always been a welcome respite to when research started to feel like real work.  

I would like to thank all my family who have always supported my absurd decision to choose science as a career and who have always been there through both the good and bad. Finally a very special thank you to Emily Keely who not only put up with my late nights behind a screen but made my time in Trinity so special.


%%% Local Variables:
%%% TeX-master: "thesis.tex"
%%% TeX-PDF-mode: t
%%% End:
